---
title: "607_Project"
author: "Raghu Ramnath"
date: "2/14/2017"
output: html_document
theme: cerulean
highlight: pygments
---

# 607 Reg Ex Project 1 

#1. Read the Tourist information text file

```{r step1, include=TRUE}
library(stringr)
getwd()
#Read the tourist info file
tour <-read.csv(("/Users/Raghu/tournamentinfo.txt"), header=F)
#Display the first 10 lines
head(tour,10)
#Display the last 10 lines
tail(tour,10)
```

#2. Create a list by removing the column names and "---" lines above and below column names.

```{r step2, include=TRUE}
tour <- tour[-c(1:4),]
head(tour)
```

#3. Read the number lines list (that has player information) separately and rating list separately.

```{r step3, include=TRUE}
player <- tour[seq(1, length(tour), 3)]
player
length(player)
rating <- tour[seq(2, length(tour), 3)]
length(rating)
```

#4. Extract the required column names by identifying the pattern from the string.

```{r step4, include=TRUE}
pairNo <- as.integer(str_extract(player, "\\d+"))  #matches the digits
Name <- str_trim(str_extract(player, "(\\w+\\s){2,3}"))   #matches the word
Points <- as.numeric(str_extract(player, "\\d+\\.\\d+"))  #matches the digits with a dot.
Opponents <- str_extract_all(str_extract_all(player, "\\d+\\|"), "\\d+") #matches digits with pipe and pick only the digits which is the id of opponent.

Won <- str_count(player, "\\Q|W  \\E")   #count of W
Loose <- str_count(player, "\\Q|L  \\E")  #count of L
Draw <- str_count(player, "\\Q|D  \\E")   #count of D

State <- str_extract(rating, "\\w+")  
Ratings <- as.integer(str_extract(str_extract(rating, "[^\\d]\\d{3,4}[^\\d]"
                                             ), "\\d+"))  # identify the string that has numbers with either 3 or 4 digits and char concatenated. pick only the first set of digits.
```

# 5. Group the required fields. calculate the mean rating. 

```{r step5, include=TRUE}
mRating <- length(player)
mRating
for (i in 1:length(player)) { 
  mRating[i] <- round(mean(Ratings[as.numeric(unlist(Opponents[pairNo[i]]))]), digits = 0);  
                            }
opData <- data.frame(Name, State, Points,Ratings,mRating,  Won, Loose, Draw);

colnames(opData) <- c("Player's Name", "State", "Total Points", "Player's Pre-Rating", " Average Pre Chess Rating of Opponents", "Won", "Lost", "Draw")
knitr::kable(opData)

```

#6. Write the data into csv file

```{r step6, include=TRUE}
   write.csv( opData, "/Users/Raghu/tour.csv")
```