---
title: "607_Project4 Document Classification"
author: "Raghu Ramnath"
date: "4/6/2017"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

The corpus for this analysis is located here: https://spamassassin.apache.org/publiccorpus/

Overview of quanteda:  1. Easy to manage texts in the form of a corpus. 2. quanteda includes tools to make it easy and fast to manuipulate the texts in a corpus, by performing the most common natural language processing tasks simply and quickly, such as tokenizing, stemming, or forming ngrams. quanteda’s functions for tokenizing texts and forming multiple tokenized documents into a document-feature matrix are both extremely fast and extremely simple to use. 3. quanteda can segment texts easily by words, paragraphs, sentences, or even user-supplied delimiters and tags.
For details refer the below link
https://cran.r-project.org/web/packages/quanteda/quanteda.pdf

# Libraries

```{r s0, include=TRUE}
#The code for this assignment requires the following R packages:

libr <- c("wordcloud","ggplot2","tm","plyr","class","stringr","stringi","RCurl","XML","SnowballC","R.utils","quanteda","class","knitr","Rweka","RTextTools")
lapply( libr, require, character.only = TRUE)
#Set Options
Options(stringsAsFactors = FALSE)
```

#Step 1: List the spam files

```{r s1, include=TRUE}

spam_file_folder <- "/Users/Raghu/spamham/spam_2/"
spam_file_names <- list.files(spam_file_folder)
#check the files
head(spam_file_names ,10)

```

#Step 2: Create functions 1.create corpus 2. Document feature matrix and 3. Data frame

```{r s2, include=TRUE}
createCorpus <- function(directory, emailType) {
    
    quantCorpus <- corpus(VCorpus(DirSource(directory = directory, encoding = "UTF-8-MAC"), 
                                    readerControl = list(language="en_US")),
                      notes=emailType)
    
    docvars(quantCorpus, "email_type") <- emailType
    docvars(quantCorpus, "source")     <- directory
    return(quantCorpus)
}

buildDFM <- function(corpus, minDoc, minCount) {
    # create the document-feature matrix(DFM)
    dfm <- dfm(corpus, ignoredFeatures = stopwords("english"), stem = TRUE)
    # word should atleast occur 10 times in 2 documents
    dfm <- dfm_trim(dfm, min_count = 10, min_docfreq = 2, sparsity = NULL, verbose = TRUE)
    return(dfm)
}

create_df_matrix <- function(dfm, emailType) {
    
    # apply the tfidf function
    #compute term frequency-inverse document frequency weights
    mat <- data.matrix(tfidf(dfm))
 
    # convert to a dataframe
    df <- as.data.frame(mat, stringsAsFactors =  FALSE)
    df$Source <- emailType
     #str(df)    
    return(df)
}

```

#Step 3: Create spam corpus

```{r s3, include=TRUE}

spamCorpus <- createCorpus("/Users/Raghu/spamham/spam_2", "spam")
summary(spamCorpus,10)

```

#Step 4: Create DFM(Document Feature Matrix) for Spam Corpus

```{r s4, include=TRUE}
dfmSpam <- buildDFM(spamCorpus, docnames(spamCorpus), 50)
dfmSpam
summary(dfmSpam,20)
dim(dfmSpam)  
topfeatures(dfmSpam, 20) 

```

#Step 5: Create Plot and word cloud for Spam

```{r s5, include=TRUE}
plot(topfeatures(dfmSpam, 100), log = "y", cex = .6, ylab = "Term frequency", main = "Top Features of Spam")

textplot_wordcloud(dfmSpam, max.words = 100,random.color = TRUE, rot.per = .25, colors = sample(colors()[2:128], 5))

```

#Step 6: Create Ham corpus

```{r s6, include=TRUE}

hamCorpus <- createCorpus("/Users/Raghu/spamham/easy_ham", "ham")
summary(hamCorpus,10)

```

#Step 7: Create DFM(Document Feature Matrix) for Ham corpus

```{r s7, include=TRUE}
dfmham <- buildDFM(hamCorpus, docnames(hamCorpus), 50)
dfmham
summary(dfmham,10)
dim(dfmham)  
topfeatures(dfmham, 20) 

```


#Step 8: Create Plot and wordcloud for Ham

```{r s8, include=TRUE}
plot(topfeatures(dfmham, 100), log = "y", cex = .6, ylab = "Term frequency", main = "Top Features of Ham")

textplot_wordcloud(dfmham, max.words = 100,random.color = TRUE, rot.per = .25, colors = sample(colors()[2:128], 5))

```

#Step 9: Stack the DFM of Spam and Ham

```{r s9, include=TRUE}
dfSpam <- create_df_matrix(dfmSpam, "spam")  
dfHam <- create_df_matrix(dfmham, "ham")  
stacked.df <- rbind.fill(dfSpam, dfHam)

#Stack the data frames of Spam and ham
# set NA values to 0
stacked.df[is.na(stacked.df)] <- 0
dim(stacked.df)
```

#Step 10: Create the tree modal

```{r s10, include=TRUE}

tdm.email <- stacked.df[, "Source"]
stacked.nl <- stacked.df[, !colnames(stacked.df) %in% "Source"]  
n <- length(tdm.email)

# taking the training size as 1000 which is one fourth of the whole size.
container <- create_container(stacked.nl,
            tdm.email, trainSize=1:1000,
            testSize=1001:n, virgin=FALSE)

slotNames(container)

#MAXENT <- train_model(container,"MAXENT")
#BOOSTING <- train_model(container,"BOOSTING")
tree_model <- train_model(container, "TREE")

#Having problems with MAXENT,SVM and BOOSTING.

#MAXENT_out <- classify_model(container, MAXENT)
#BOOSTING_out <- classify_model(container, BOOSTING)
tree_model_out <- classify_model(container, tree_model)

head(tree_model_out,5)

labels_out <- data.frame(
  correct_label = tdm.email[1001:n],
  tree = as.character(tree_model_out[,1]),
  stringAsFactors = F)

#ensure the dataframe does not have factor columns
labels_out$tree <- as.character(labels_out$tree)

#TREE Performance
table(labels_out[,1] == labels_out[,2])
```

#Step 11: Create KNN modal (k-Nearest Neighbor)

```{r s11, include=TRUE}
## Create training and test datasets 
train.idx <- sample(nrow(stacked.df), ceiling(nrow(stacked.df) * 0.7))
test.idx <- (1:nrow(stacked.df)) [-train.idx]
head(train.idx,5)
head(test.idx,5)
length(train.idx)  
length(test.idx)  

#kNN prediction using the training and test datasets
knn.pred <- knn(stacked.nl[train.idx, ], stacked.nl[test.idx, ], tdm.email[train.idx])
conf.mat <- table("Predictions" = knn.pred, Actual = tdm.email[test.idx])
conf.mat
#df.pred <- cbind(knn.pred, stacked.nl[test.idx, ])
#head(df.pred,5)

```

#Conclusion:
Comparing the two modals- KNN and Tree, I could see that the predictions match. 397 Spam reported by these two modals.