---
title: "Raghu_607_wk3_DP"
author: "Raghu Ramnath"
date: "2/14/2017"
output: html_document
highlight: pygments
theme: cerulean
---

#607 Week3 Assignment
Problems 3 and 4 from chapter 8 of Automated Data Collection in R.

Problem3: Here is the referenced code for the introductory example in #3:

3. Copy the introductory example. The vector name stores the extracted names.

```{r s3, include=TRUE}
#Load package
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"   

raw.data

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))

name
phone
#phone number is in the format XXX-XXX-XXXX
```

3.1. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r s3.1, include=TRUE}
# remove initials
name2 <- sub(" [A-z]{1}\\. "," ",name)  
name2
# switch last,first to first last
name3 <- sub("(\\w+),\\s(\\w+)","\\2 \\1", name2)  
# two names has "," which need to be switched.
name3
# remove titles
name4 <- sub("[A-z]{2,3}\\. ","",name3)  
name4
```

3.2. Construct a logical vector indicating whether a character has a title.

```{r s3.2, include=TRUE}

title_yn <- str_detect(name3,"[A-z]{2,3}\\. ")
df <- data.frame(name3,title_yn)
df
# two of the name has title of which one is 2 char and the other is 3 char
```

3.3. Construct a logical vector indicating whether a character has a second name

```{r s3.3, include=TRUE}
second_name <- str_detect(name," [A-z]{1}\\. ")
df <- data.frame(name,second_name)
df
# One name has a middle name.
```
#Problem 4

4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
4.1. [0-9]+\\$
```{r s4.1, include=TRUE}
#This expression describes string of one or more digits followed by dollar sign.
pattern="[0-9]+\\$"
examples=c("123$43","xyz123$9","XYZ$")
str_detect(examples,pattern)
```

4.2. \\b[a-z]{1,4}\\b
```{r s4.2, include=TRUE}
#This expression describes string of one to four lower case letters.
words="\\b[a-z]{1,4}\\b"
examples=c("piG2","CAT","dogs")
str_detect(examples,words)
```

4.3. .*?\\.txt$
```{r s4.3, include=TRUE}
#This expression describes a string ending in “.txt”.
pattern=".*?\\.txt$"
examples=c(".txt","test.txt","abc/123.txt","a$b#1.txt")
str_detect(examples,pattern)

```

4.4. \\d{2}/\\d{2}/\\d{4}
```{r s4.4, include=TRUE}
#This expression describes 2 digits followed by a forward slash, 2 more digits, another forward slash, and four more digits like a date format 02/14/2017
pattern="\\d{2}/\\d{2}/\\d{4}"
examples=c("02/14/2017","DD11/12MM/2020RR")
str_detect(examples,pattern)

```

4.5. <(.+?)>.+?</\\1>
```{r s4.5, include=TRUE}
#This expression describes XML tags.
pattern="<(.+?)>.+?</\\1>"
examples=c("<tag>Text</tag>","<tag>123</tag>")
str_detect(examples,pattern)
```
#Problem 9

9. The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

```{r s9, include=TRUE}
hidden <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

unlist(str_extract_all(hidden, "[A-Z]"))

unlist(str_extract_all(hidden, "[a-z]"))

unlist(str_extract_all(hidden, "[0-9]"))

#The hidden message is revealed by extracting just the uppercase characters
```